---
title: "Dataset Cleaning — Capture Records"
format: html
editor: visual
---

## Setup

Set the paths to your input Excel file and desired output CSV before running.

```{r}
#| label: setup
#| message: false

library(tidyverse)
library(readxl)

df <- read_excel("../data_clean/Small_Mammals_75-01.xlsx")
df <- read_csv("../data_clean/Small_Mammals_75-01.csv")
```

------------------------------------------------------------------------

## Flag Invalid Mark Values

Valid marks must be exactly one of the two-letter codes (`LF`, `RF`, `RR`, `LR`, `UM`, `ST`) followed by a single digit (e.g. `LF3`, `ST0`).

```{r}
#| label: flag-mark

valid_mark_pattern <- "^(LF|RF|RR|LR|UM|ST)[0-9]$"

df <- df |>
  mutate(
    mark_flag = if_else(
      !str_detect(Mark, valid_mark_pattern),
      "invalid Mark format/combo",
      NA_character_
    )
  )
```

------------------------------------------------------------------------

## Flag Recapture Issues

Two conditions are flagged:

1.  **Invalid value** — anything other than `"Y"` or `"N"`
2.  **Order mismatch** — within each unique combination of Plot, Location, Genus, Species_Captured, Sex, and Mark, the first occurrence should be `"N"` and all subsequent should be `"Y"`

```{r}
#| label: flag-recapture

df <- df |>
  mutate(Trapping_Day = as.integer(Trapping_Day),
         Trapping_Month = as.integer(Trapping_Month),
         Plot = as.integer(Plot)) |> 
  arrange(Location, Plot, Genus, Species_Captured, Sex, Mark, Trapping_Month, Trapping_Day) |> 
  group_by(Location, Plot, Genus, Species_Captured, Sex, Mark) |>
  mutate(expected_recapture = if_else(row_number() == 1L, "N", "Y")) |>
  ungroup() |>
  mutate(
    recapture_flag = case_when(
      !Recapture %in% c("Y", "N") ~ paste0("invalid value (not Y/N): '", Recapture, "'"),
      Recapture != expected_recapture ~ paste0(
        "expected ", expected_recapture, ", got ", Recapture
      ),
      .default = NA_character_
    )
  )
```

------------------------------------------------------------------------

## Summary of Flagged Rows

Review this table before running the interactive session below.

```{r}
#| label: flag-summary

flagged <- df |>
  filter(!is.na(mark_flag) | !is.na(recapture_flag))

n_flagged <- nrow(flagged)

if (n_flagged == 0) {
  cat("No issues found — dataset looks clean!\n")
} else {
  cat(sprintf("%d row(s) flagged for review.\n\n", n_flagged))

  flagged |>
    mutate(.row = which(!is.na(df$mark_flag) | !is.na(df$recapture_flag))) |>
    select(
      .row, Plot, Location, Genus, Species_Captured, Sex,
      Mark, mark_flag,
      Recapture, recapture_flag
    ) |>
    print(n = Inf)
}
```

------------------------------------------------------------------------

## Interactive Review

Run the two chunks below **from an R console or RStudio** (not during rendering). They use `readline()` to walk you through each flagged row and collect corrections.

```{r}
#| label: review-function
#| eval: false

review_flags <- function(df) {
  flagged_idx <- which(!is.na(df$mark_flag) | !is.na(df$recapture_flag))

  if (length(flagged_idx) == 0) {
    message("No flagged rows — nothing to review.")
    return(df |> select(-mark_flag, -recapture_flag, -expected_recapture))
  }

  cat(sprintf("\n%d row(s) flagged for review.\n", length(flagged_idx)))

  for (i in flagged_idx) {
    row <- df[i, ]

    cat("\n\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\n")
    cat(sprintf(
      "Row %-4d | Plot: %-6s | Location: %-6s | Genus: %-14s | Species: %-16s | Sex: %s\n",
      i, row$Plot, row$Location, row$Genus, row$Species_Captured, row$Sex
    ))

    # --- Mark ---
    if (!is.na(row$mark_flag)) {
      cat(sprintf("  Mark:      %-10s  \u2190 %s\n", paste0("'", row$Mark, "'"), row$mark_flag))
      new_val <- readline(sprintf("  New Mark value       (Enter to keep '%s'): ", row$Mark))
      if (nchar(trimws(new_val)) > 0) {
        df$Mark[i] <- trimws(new_val)
        cat(sprintf("  \u2713 Mark updated to '%s'\n", trimws(new_val)))
      }
    }

    # --- Recapture ---
    if (!is.na(row$recapture_flag)) {
      cat(sprintf("  Recapture: %-10s  \u2190 %s\n", paste0("'", row$Recapture, "'"), row$recapture_flag))
      new_val <- readline(sprintf("  New Recapture value  (Enter to keep '%s'): ", row$Recapture))
      if (nchar(trimws(new_val)) > 0) {
        df$Recapture[i] <- trimws(new_val)
        cat(sprintf("  \u2713 Recapture updated to '%s'\n", trimws(new_val)))
      }
    }
  }

  cat("\n\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\u2500\n")
  cat("Review complete.\n")

  df |> select(-mark_flag, -recapture_flag, -expected_recapture)
}
```

```{r}
#| label: run-review-and-save
#| eval: false

# Run this in your R console after sourcing the chunk above.
# It will walk you through each flagged row, then save the cleaned file.

df_clean <- review_flags(df)

```

Ran this and completed all rows, just have some notes

1.  Need to confirm that all "weird marks" (e.g. LF12, RF1RR1) are explicit in the binder

```{r}
# Save clean data
write_csv(df_clean, "../data_clean/Small_Mammals_75-01.csv")
cat(sprintf("Saved cleaned dataset to: %s\n", df_clean))
```

```{r}
df_clean[7, "Mark"] <- "UM1"

df_clean[70, "Mark"] <- "RF2"

df_clean[79, "Mark"] <- "LF1"
```
